---
title: "Data Cleaning of OCTGN Netrunner dataset"
author: "Stephan Koenig"
date: "August 8, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Goal
=====

The provided code prepares the latest [Netrunner OCTGN dataset][1] provided by
[db0][2] for further analysis.

### Helper functions

```{r helper_functions, echo=FALSE}
load_packages <- function(required_packages) {
    # Takes character or character vector of required_packages, installs any
    # missing packages and only attaches those that are not yet attached.
    
    # Check which packages are not installed.
    missing_packages <- !is.element(required_packages, installed.packages()[,1])
    missing_packages <- required_packages[missing_packages]
    
    # Install missing packages.
    if (length(missing_packages) != 0) {
        install.packages(missing_packages)
        packages <- paste(missing_packages, sep = ", ")
        system_message(paste("Package(s)", packages, "installed."))
    }
    
    # Check which packages are already loaded.
    unattached_packages <- !is.element(required_packages, (.packages()))
    attached_packages <- required_packages[!unattached_packages]
    unattached_packages <- required_packages[unattached_packages]
    
    # Attach any unattached packages.
    if (length(unattached_packages != 0)) {
        lapply(unattached_packages, library, character.only = TRUE)
        packages <- paste(unattached_packages, sep = ", ")
        system_message(paste("Package(s)", packages, "attached."))
    }
    
    if (length(attached_packages != 0)) {
        packages <- paste(attached_packages, sep = ", ")
        system_message(paste("Package(s)", packages, "already attached."))
    }
}
    
system_message <- function(text) {
    # Generates system output without quotes.
    
    message(noquote(text))
}
```

Data Cleaning
=====

In general, ordered factor levels are assigned to ensure a consitent layout
for graphs.

### Loading Dataset

```{r load_dataset}
load_packages("dplyr")
# Dataset file must be in working directory.
filename <- "OCTGN_stats_anonymized-2015-02-20.csv"
nrdata <- tbl_df(read.csv(filename))
```

### Change of variable names

Rename variables to consistent convention. Reformat game_start as POSIXct,
convert corp deck size to integer.

```{r renaming_and_drops}
# Rename variables that are kept.
load_packages("lubridate")
nrdata <- rename(nrdata,
                 corp_player_id = Corp_Player,
                 runner_player_id = Runner_Player,
                 game_start = GameStart,
                 turns_played = Turns_Played,
                 version = Version,
                 corp_inf = Corp_Inf,
                 runner_inf = Runner_Inf,
                 corp_score = Corp_Score,
                 runner_score = Runner_Score,
                 runner_deck_size = Runner_Deck_Size) %>%
    
                  # Convert POSIXct.
           mutate(game_start = ymd_hms(game_start),
                  # Only keep corp deck sizes legal size 40 or larger.
                  corp_deck_size = as.integer(Corp_Deck_Size))
```

### Corp and Runner

Assign the commonly used ID names (mostly card title) as variable and
separately a  faction variable. Determine any decks that violate inf or 
deck size limits.

```{r cleaning_corp}
# Load lookup table for corps (manually curated) and match to OCTGN names.
lookup_corp <- tbl_df(read.csv("lookup_corp.csv"))
match_corps <- factor(match(nrdata$Corp_Faction, lookup_corp$octgn_id))

# Look up corp id and faction and order by sequence given in 
nrdata$corp_id <- factor(lookup_corp$id[match_corps],
                         levels = lookup_corp$id,
                         ordered = TRUE)
                         
nrdata$corp_faction <- factor(lookup_corp$faction[match_corps],
                              levels = unique(lookup_corp$faction),
                              ordered = TRUE)

# Look up deck size and inf limits for corp id and determine if decks are legal.
nrdata$corp_legal_ds <- lookup_corp$deck_size[match_corps]
nrdata$corp_legal_ds <- nrdata$corp_deck_size >= nrdata$corp_legal_ds

nrdata$corp_legal_inf <- lookup_corp$inf[match_corps]
nrdata$corp_legal_inf <- nrdata$corp_inf <= nrdata$corp_legal_inf

rm(lookup_corp, match_corps)

# List changed names.
nrdata[c("Corp_Faction", "corp_faction", "corp_id")]

# List decks below legal size.
nrdata[!is.na(nrdata$corp_legal_ds) & !nrdata$corp_legal_ds,
       c("corp_id", "corp_deck_size", "corp_legal_ds")]

# List decks with too much inf.
nrdata[!is.na(nrdata$corp_legal_inf) & !nrdata$corp_legal_inf,
       c("corp_id", "corp_inf", "corp_legal_inf")]
```

```{r cleaning_runner}
# Load lookup table for runners (manually curated) and match to OCTGN names.
lookup_runner <- tbl_df(read.csv("lookup_runner.csv"))
match_runners <- factor(match(nrdata$Runner_Faction, lookup_runner$octgn_id))

# Look up runner id and faction and order by sequence given in lookup table.
nrdata$runner_id <- factor(lookup_runner$id[match_runners],
                           levels = lookup_runner$id,
                           ordered = TRUE)
                         
nrdata$runner_faction <- factor(lookup_runner$faction[match_runners],
                                levels = unique(lookup_runner$faction),
                                ordered = TRUE)

# Look up deck size and inf limits for runner id and determine if decks are legal.
nrdata$runner_legal_ds <- lookup_runner$deck_size[match_runners]
nrdata$runner_legal_ds <- nrdata$runner_deck_size >= nrdata$runner_legal_ds

nrdata$runner_legal_inf <- lookup_runner$inf[match_runners]
nrdata$runner_legal_inf <- nrdata$runner_inf <= nrdata$runner_legal_inf

rm(lookup_runner, match_runners)

# List changed names.
nrdata[c("Runner_Faction", "runner_faction", "runner_id")]

# List decks below legal size.
nrdata[!is.na(nrdata$runner_legal_ds) & !nrdata$runner_legal_ds,
       c("runner_id", "runner_deck_size", "runner_legal_ds")]

# List decks with too much inf.
nrdata[!is.na(nrdata$runner_legal_inf) & !nrdata$runner_legal_inf,
       c("runner_id", "runner_inf", "runner_legal_inf")]
```

### Add Release and Cycle

Add information on what cards were most likely supported at the time of the
game to capture changes in the meta due to new releases.

```{r release_cycle}
# Set any empty string in version to NA and sort.
load_packages("naturalsort")

nrdata$version[nrdata$version == ""] <- NA
nrdata$version <- ordered(nrdata$version,
                         levels = naturalsort(unique(nrdata$version)))

# Load lookup table for conversion of OCTGN version number to release.
lookup_octgn <- read.csv("lookup_octgn.csv")

# Convert version to release.
nrdata$release <- match(nrdata$version, lookup_octgn$version)
nrdata$release <- factor(lookup_octgn$release[nrdata$version],
                         levels = unique(lookup_octgn$release),
                         ordered = TRUE)

# Load lookup table for conversion of release to cycle.
lookup_release <- read.csv("lookup_release.csv")

# Convert release to cycle.
nrdata$cycle <- match(nrdata$release, lookup_release$release)
nrdata$cycle <- factor(lookup_release$cycle[nrdata$cycle],
                       levels = unique(lookup_release$cycle),
                       ordered = TRUE)

rm(lookup_octgn, lookup_release)

# Show added data.
unique(nrdata[!is.na(nrdata$version) & nrdata$cycle == "Lunar",
              c("version", "release", "cycle")])
```

### Match Result

Convert result variable into one of four win types ("agenda", "concede",
"decked", "flatlined") and a winning side.

```{r result}
# Load lookup table for conversion of result.
lookup_result <- read.csv("lookup_result.csv")

# Convert result into win type and winner.
nrdata$winner <- nrdata$win_type <- match(nrdata$Result, lookup_result$result)
nrdata$winner <- factor(lookup_result$winner[nrdata$winner],
                        levels = unique(lookup_result$winner),
                        ordered = TRUE)
nrdata$win_type <- factor(lookup_result$type[nrdata$win_type],
                          levels = unique(lookup_result$type),
                          ordered = TRUE)

rm(lookup_result)

# Show reformatted results.
unique(nrdata[c("Result", "winner", "win_type")])
```

### Subset legal games

```{r illegal_games}
system_message(paste("Length complete data set:", dim(nrdata)[1]))

# Only keep games with regular duration.
legal <- nrdata$Duration %in% 0:120

# Only keep games within regular score range.
legal <- legal &
         nrdata$corp_score %in% 0:12 &
         nrdata$runner_score %in% -3:24

# Only keep games with legal decks.
legal <- legal &
         nrdata$corp_legal_ds & nrdata$corp_legal_inf &
         nrdata$runner_legal_ds & nrdata$runner_legal_inf

nrdata <- nrdata[legal, ]

rm(legal)

system_message(paste("Length legal games data set:", dim(nrdata)[1]))
```

### Final variable selection

Only relevant variables for plotting are kept and the final data set is saved
as an R object.

```{r variable_subsetting}
nrdata <- select(nrdata,
                 game_start,
                 version,
                 cycle,
                 release,
                 corp_player_id,
                 corp_faction,
                 corp_id,
                 corp_deck_size,
                 corp_inf,
                 runner_player_id,
                 runner_faction,
                 runner_id,
                 runner_deck_size,
                 runner_inf,
                 turns_played,
                 winner,
                 win_type,
                 corp_score,
                 runner_score)

save(nrdata, file = "nrdata.RData")
```


[1]: https://docs.google.com/uc?export=download&confirm=qg-_&id=0B-gMiPlH3rBAcmZmVnhNLU5wVEE
[2]: http://dbzer0.com/